---
title: "Tidyverse Assignment - Ramen Ratings"
author: "Stephen Haslett"
date: "11/25/2019"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment Overview

**Create an Example** Using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset.

```{r, eval=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(rmdformats)
```


## Assignment Response

For this assignment, I have chosen to demonstrate the basic functionailty supplied by the dplry package that comes with tidyverse.

## Datasource
https://www.kaggle.com/residentmario/ramen-ratings/data


#### Import the Ramen Ratings dataset using the "readr" package.


```{r, eval=TRUE, message=FALSE, warning=FALSE}
ramen_ratings <- read.csv("https://raw.githubusercontent.com/stephen-haslett/FALL2019TIDYVERSE/master/ramen-ratings.csv", header = TRUE)
```


#### Check that the data imported successfully.

```{r, eval=TRUE, message=FALSE, warning=FALSE}
ramen_ratings <- as_tibble(ramen_ratings)
head(ramen_ratings)
```


#### Using the "dplyr" package's "slice" function, select a specific range of rows from the dataset.

```{r, eval=TRUE, message=FALSE, warning=FALSE}
# Select rows 10 to 20 from the dataset.
selected_rows <- slice(ramen_ratings, 10:19)
selected_rows
```


#### Now use the "slice" function to select 2 row ranges, and 4 specific rows.

```{r, eval=TRUE, message=FALSE, warning=FALSE}
# Select rows 2 to 9, 26 to 30, and rows 40, 21, 16, and 35.
multiple_selected_rows <- ramen_ratings %>% slice(c(2:9, 26:30, 40, 21, 16, 35))
multiple_selected_rows
```


#### Use dplyr's "filter" function to select rows based on specified conditions.

_In this example, we will only select rows that have a 5 Star rating._

```{r, eval=TRUE, message=FALSE, warning=FALSE}
five_stars <- filter(ramen_ratings, Stars == 5)
five_stars
```


#### Finally, use dplyr's "select" function to select specific data columns from the dataset.

_In this example, we will select the "Brand", "Variety", "Country", "Stars", and "Top.Ten" columns for brands that have a Top.Ten listing._

```{r, eval=TRUE, message=FALSE, warning=FALSE}
# Select only the rows with Top.Ten column entries whilst removing those
# with NA, empty, or "\n" values.
ramen_ratings_filtered <- filter(ramen_ratings, Top.Ten != "" & !is.na(Top.Ten) & Top.Ten != "\n")

ramen_top_ten <- select(ramen_ratings_filtered, c("Brand", "Variety", "Country", "Stars", "Top.Ten"))
ramen_top_ten
```



## Devanshu Mehrotra additions based on task 2 of the assignment

### Addition 1: We will be using the ```group_by``` , ```tally``` and top_n functions to create a dataFrame containing countries along with the number of ramen restaurants reviewed.

According to the dplyr documentation the group_by() function takes an existing tbl and converts it into a grouped tbl where operations are performed "by group". Tally is just a wrapper to summarize the data by counting how many elements within the group. We will also use the top_n function to seperate the top 10 countries by count of ramen establishments


```{r, Groupby and Tally}

# group by country and name of the column in which the tally is to be stored. Pick the top 10 countries by count
 ramen_counts <- (ramen_ratings) %>%
  group_by(Country) %>%
  tally(name = 'count') %>% 
  top_n(10)

```


### Addition 2: We will use ggplot() to visualize the datadrame from addition 1 above. This will happen in three steps

1. ggplot() initialize a ggplot object which declares input data frame for a graphic and sets aesthetics intended to be common throughout all subsequent layers. Syntax for creating a ggplot object is
ggplot(data = NULL, mapping = aes(), ...)

Where: - data- dataset to used for plotting.

mapping - list of aesthetic mappings to use for plotting

Note that fill=Country just tells ggplot to color each country seperately

2. geom_bar() creates a bar plot where the argument stat='identity' ensures that the length of the bars are determined by the counts of y axis 

3. Other arguments to determine the overall look of the visualiation


```{r, ratings}
ggplot(ramen_counts, aes(x=Country, y=count, fill=Country ))+geom_bar(stat = "identity" ) + theme(axis.text.x = element_text(size=12, angle = 90, hjust = 1))

```


